---
title: "KNN Regression"
output: html_notebook
---
First, some codes about how to split the data into training, validation and test.
1.Create a data frame
```{r}
data <- read.table("GradedHW1-All-Data.csv", header = T, sep = ",")
```
2.Remove all observations except those that are single family homes.
```{r}
data <- data[data$Bldg.Type == "1Fam",]
```
3.Split the data frame into training, validation, and test data sets.
```{r}
train_size <- ceiling(0.5*nrow(data))
set.seed(12)
train_data <- sample(nrow(data), size = train_size)

last <- data[-train_data,]
vali_size <- ceiling(0.5*nrow(last))
set.seed(12)
vali_data <- sample(nrow(last), size = vali_size)

train <- data[train_data,]
vali <- last[vali_data,]
test <- last[-vali_data,]
```


1.Import data
```{r}
data_train <- read.table("TrainData.csv", sep = ",", header = T)
data_test <- read.table("TestData.csv", sep = ",", header = T)
data_vali <- read.table("ValidationData.csv", sep = ",", header = T)

data_train[c(1:4,6:19,21:38,40:46,48:49,51,53:78)] <- list(NULL)
data_vali[c(1:4,6:19,21:38,40:46,48:49,51,53:78)] <- list(NULL)
data_test[c(1:4,6:19,21:38,40:46,48:49,51,53:78)] <- list(NULL)

data_train$Age <- 2017- data_train$Year.Built
data_vali$Age <- 2017- data_vali$Year.Built
data_test$Age <- 2017- data_test$Year.Built

data_train$Year.Built <- NULL
data_vali$Year.Built <- NULL
data_test$Year.Built <- NULL
```
2.Examine the variables
```{r}
head(data_train)
str(data_train)
summary(data_train)
```


Compare KNN regression with standardizing variables and without standardizing variables

1.knn-regression with standardizing variables
a)Standardize the variables
```{r}
train <- data_train[complete.cases(data_train),]
test <- data_test[complete.cases(data_test),]
vali <- data_vali[complete.cases(data_vali),]

standardize <- function(x) {
  return( ( x - min(x) ) / ( max(x) - min(x) ) )
}

Tr <- as.data.frame(lapply(train[,c(1,2,3,4,5,7)], standardize))
Te <- as.data.frame(lapply(test[,c(1,2,3,4,5,7)], standardize))
Va <- as.data.frame(lapply(vali[,c(1,2,3,4,5,7)], standardize))

summary(Tr)
```
b)Find the best K through validation data
```{r}
m1 <- knn.reg(train=Tr, y=SalePrice, test= Va, k=1)
y_p <- m1$pred
MSE <- mean((y_p- vali$SalePrice)^2)

kbest <- 0
klist <- seq(nrow(Va))
for (k in klist) {
  m1 <- knn.reg(train=Tr, y=SalePrice, test= Va, k)
  y_p <- m1$pred
  mse.test <- mean((y_p- vali$SalePrice)^2)
  if (mse.test < MSE) {
    MSE <- mse.test
    kbest <- k
  } 
}
```
c)Model's generalization performance
```{r}
m2 <- knn.reg(train=Tr, y=SalePrice, test= Te, k=kbest)
y_pred <- m2$pred
MSE.test <- mean((y_pred- test$SalePrice)^2)
sqrt(MSE.test)
```


2.knn-regression without standardizing variables

```{r}
Tr <- train[,c(1,2,3,4,5,7)]
Te <- test[,c(1,2,3,4,5,7)]
Va <- vali[,c(1,2,3,4,5,7)]

summary(Tr)

m1 <- knn.reg(train=Tr, y=SalePrice, test= Va, k=1)
y_p <- m1$pred
MSE <- mean((y_p- vali$SalePrice)^2)

kbest <- 0
klist <- seq(nrow(Va))
for (k in klist) {
  m1 <- knn.reg(train=Tr, y=SalePrice, test= Va, k)
  y_p <- m1$pred
  mse.test <- mean((y_p- vali$SalePrice)^2)
  if (mse.test < MSE) {
    MSE <- mse.test
    kbest <- k
  } 
}

m2 <- knn.reg(train=Tr, y=SalePrice, test= Te, k=kbest)
y_pred <- m2$pred
MSE.test <- mean((y_pred- test$SalePrice)^2)
sqrt(MSE.test)

```


Build KNN regression Model:

1.Import the data 
```{r}
data_train <- read.table("GradedHW1-Train-Data.csv", sep = ",", header = T)
data_test <- read.table("GradedHW1-Test-Data.csv", sep = ",", header = T)
data_vali <- read.table("GradedHW1-Validation-Data.csv", sep = ",", header = T)

data_train[c(1:4,6:19,21:38,40:46,48:49,51,53:78)] <- list(NULL)
data_vali[c(1:4,6:19,21:38,40:46,48:49,51,53:78)] <- list(NULL)
data_test[c(1:4,6:19,21:38,40:46,48:49,51,53:78)] <- list(NULL)

data_train$Age <- 2017- data_train$Year.Built
data_vali$Age <- 2017- data_vali$Year.Built
data_test$Age <- 2017- data_test$Year.Built

data_train$Year.Built <- NULL
data_vali$Year.Built <- NULL
data_test$Year.Built <- NULL

train <- data_train[complete.cases(data_train),]
test <- data_test[complete.cases(data_test),]
vali <- data_vali[complete.cases(data_vali),]

```


2.Transform variables
```{r}
plot(train)
```
a)Not indepedent
```{r}
plot(train$Total.Bsmt.SF, train$Gr.Liv.Area)
cor(train$Total.Bsmt.SF, train$Gr.Liv.Area)

train$sTotal <- sqrt(train$Gr.Liv.Area+train$Total.Bsmt.SF)
plot(train$sTotal, train$SalePrice)
cor(train$sTotal, train$SalePrice)#0.78
test$sTotal <- sqrt(test$Gr.Liv.Area+test$Total.Bsmt.SF)
vali$sTotal <- sqrt(vali$Gr.Liv.Area+vali$Total.Bsmt.SF)
```
b)Non-linear relationship
```{r}
train$logage <- log(train$Age)
cor(train$logage, train$SalePrice)
test$logage <- log(test$Age)
vali$logage <- log(vali$Age)
```
c)Heteroscedasticity
```{r}
train$loglot <- log(train$Lot.Area)
test$loglot <- log(test$Lot.Area)
vali$loglot <- log(vali$Lot.Area)

train$lot1 <- train$loglot*train$Full.Bath 
cor(train$lot1, train$SalePrice) #0.62
plot(train$lot1, train$SalePrice)
test$lot1 <- test$loglot*test$Full.Bath
vali$lot1 <- vali$loglot*vali$Full.Bath

train$lot2 <- train$loglot*train$sTotal
cor(train$lot2, train$SalePrice) #0.77
test$lot2 <- test$loglot*test$sTotal
vali$lot2 <- vali$loglot*vali$sTotal
```
d)Counts
```{r}
train$bed2 <- train$Bedroom.AbvGr^2
cor(train$Bedroom.AbvGr, train$SalePrice)
cor(train$bed2, train$SalePrice)
test$bed2 <- test$Bedroom.AbvGr^2
vali$bed2 <- vali$Bedroom.AbvGr^2

train$bath2 <- train$Full.Bath^2
cor(train$bath2, train$SalePrice)
cor(train$Full.Bath, train$SalePrice)
test$bath2 <- test$Full.Bath^2
vali$bath2 <- vali$Full.Bath^2
```


3.Build the knn model without standardizing variables
```{r}
Tr <- train[,c(8,9,11,13)]
Te <- test[,c(8,9,11,13)]
Va <- vali[,c(8,9,11,13)]

summary(Tr)
```
a)choose the best K
```{r}
m1 <- knn.reg(train=Tr, y=train$SalePrice, test= Va, k=1)
y_p <- m1$pred
MSE <- mean((y_p- vali$SalePrice)^2)
kbest <- 0
klist <- seq(nrow(Va))
for (k in klist) {
  m1 <- knn.reg(train=Tr, y=train$SalePrice, test= Va, k)
  y_p <- m1$pred
  mse.test <- mean((y_p- vali$SalePrice)^2)
  if (mse.test < MSE) {
    MSE <- mse.test
    kbest <- k
  } 
}
```
b)Final MSE
```{r}
m2 <- knn.reg(train=Tr, y=SalePrice, test= Te, k=kbest)
y_pred <- m2$pred
MSE.test <- mean((y_pred- test$SalePrice)^2)
sqrt(MSE.test)
```








